---
title: "Ecological Footprints"
output: html_notebook
---

##Introduction
In our discussion of ecological economics, we are examining how many earths it takes to sustain the human population.  Students were asked to go to [https://www.footprintcalculator.org/](https://www.footprintcalculator.org/) and calculate their ecological footprint for home and for school.

Here we will make some plots of the data

###Set up the workspace
```{r}
rm(list=ls())
library(tidyverse)
library(lubridate)
library(ggthemes)
library(here)
```
###Pull in some data
```{r}
Feet<-read.csv(here("Data", "footprints.csv"), stringsAsFactors = T)
```
###Clean up the data

```{r}
Feet$Year<-factor(Feet$Year, levels = c("2018", "2019", "2021", "2022", "2023"))
Feet$Highest.consumption.category<-as.factor(Feet$Highest.consumption.category)

levels(Feet$Location)
levels(Feet$Highest.consumption.category)
fix<-which(Feet$Highest.consumption.category == "Mobilty")
Feet$Highest.consumption.category[fix]<-"Mobility"
fix<-which(Feet$Highest.consumption.category == "Shelter ")
Feet$Highest.consumption.category[fix]<-"Shelter"
fix<-which(Feet$Highest.consumption.category == "")
Feet$Highest.consumption.category[fix]<-NA
Feet$Highest.consumption.category<-factor(Feet$Highest.consumption.category)
levels(Feet$Highest.consumption.category)
```

###Make some plots

First look at campus vs home
```{r}
ggplot(Feet, aes(Year, Number.of.earths, fill = Location, color = Location))+
     geom_boxplot(aes(outlier.color = "black", outlier.size = 3), color = "black")+
	scale_fill_manual(values = c("white", "gray"))+
	geom_point(aes(group=Location, fill = Location, color = Location),  position=position_dodge(width=0.75), color = "black")+
	#scale_color_manual(values = c("white", "gray"))+
	coord_flip()+
	geom_hline(yintercept = 1, color = "red", size = 1)+
	labs(x = "Year", y = "Mean number of Earths")+
	ylim(0,8)+
	theme_fivethirtyeight()
	


	

```
Now save
```{r}
ggsave (here("Code", "Footprints", "earths_by_year.png"),width = 6, height = 4, device = "png")
```


Now see if we can plot the overshoot dates by year and location


Prep the data
```{r}
Feet$Overshoot<-as.Date(Feet$Overshoot,"%m/%d/%y" )
```
Now see if I can plot julian dates

```{r}
Feet$Julian<-format(Feet$Overshoot, "%j")
Feet$Julian<-as.numeric(Feet$Julian)
```

```{r}
ggplot(Feet, aes(Julian, color = Year, fill = Year))+
     geom_histogram()+
	labs(x = "Julian date", y = "Number of people", title = "Histogram of overshoot dates")+
	theme_fivethirtyeight()
ggsave("Overshoot_dates.png", device = "png")
```


Now see if I can summarize and add some means to the plots

```{r}
NewFeet<-Feet %>% group_by (Year, Place) %>% summarize (
	MeanE = mean(Earths, na.rm = T),
	semE = sd(Earths)/sqrt(sum(!is.na(Earths))),
	meanJul = mean(Julian, na.rm = T),
	semJul =  sd(Julian)/sqrt(sum(!is.na(Julian)))
)
```
Now plot that

```{r}
ggplot(NewFeet, aes(Year, MeanE,  fill = Place))+
	geom_col(position = "dodge", width = 0.8, color = "black")+
	scale_fill_manual(values= c("white", "gray"))+
	geom_errorbar(aes(ymin= MeanE - semE,
				   ymax = MeanE+semE), width = 0.2, position = position_dodge(width = 0.8))+
	geom_hline(yintercept = 1, color = "red", size = 1)+
	labs(x = "Year", y = "Mean number of Earths", title = "Mean footprint in number of earths")+
	coord_flip()+
	theme_fivethirtyeight()
```
Now save
```{r}
ggsave("MeanEarthsvsYear.png", device = "png")
```
Between 2019 and 2021, the footprint calculator started providing more information, so there is some we can only look at from the 2021 data (and I forgot to ask them to estimate home vs. campus in 2021).

Let's start by renaming some columns
```{r}
names(F2021)<- c("Name", "Overshoot", "Earths","Footprint_Ha", "TonsCO2", "Carbon_percent", "Category")
```
Now deal with the fact that the category variable came in with mis-labeled factors
```{r}
levels(F2021$Category)
fix<-which(F2021$Category == "mobility"|F2021$Category == "Mobility"|F2021$Category== "Mobilty")
F2021$Category[fix]<-"Mobility"
F2021$Category<-factor(F2021$Category)
levels(F2021$Category)
```
Let's get a histogram of main causes of footprint

```{r}
ggplot(F2021, aes(Category))+
	geom_histogram(stat = "count")+
	labs(title = "Histogram of footprint factors", x = "Factor", y = "Number of responses")+
	theme_fivethirtyeight()
ggsave("MajorCauses.png", device = "png")
```

Now I'd like to reshape the DF for plotting purposes so that I have an index for the variables in Earths:Carbon_Percent (the numeric ones). Use tidyr

```{r}
LF2021<-F2021 %>% pivot_longer(Earths:Carbon_percent, names_to = "Variable")
```
Now summarize the data
```{r}
LF_summary<-LF2021 %>% group_by(Variable) %>% summarise(
	avg_val=mean(value, na.rm = T),
	sd_val = sd(value, na.rm = T)
)
```
Now extract carbon data
```{r}
Carbon<-LF2021 %>% filter(Variable == "TonsCO2" | Variable == "Carbon_percent")
```
And make Overshoot into a date
```{r}
Carbon$Overshoot<-as.character(Carbon$Overshoot)
Carbon$Overshoot<-dmy(Carbon$Overshoot)
```


Make some plots
```{r}
ggplot(F2021, aes(Earths))+
	geom_histogram()
```









