---
title: "Biogeography"
author: "Erika Barthelmess"
date: "2/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Step 1. Set up the R workspace

```{r, echo = FALSE}
rm(list = ls()) #clears the workspace for a fresh start
library(dplyr) #turns on dplyr package
library(ggplot2) #turns on ggplot2 package
library(here) #helps with file paths
```

### Step 2. Import and organize Data

#### A.
Open the file that has Rhat over time

```{r}
library(readr)
Rhat <- read.csv(here("Data", "Biogeography", "R_hat over time.csv"), stringsAsFactors = T)

str(Rhat) #presents how the data frame is structured
```
#### B. 
Summarize the data a bit

Begin by controlling variable types a bit
```{r}
Rhat$Replicate<-as.factor(Rhat$Replicate)
```

#### C. 
Now fill in the values based on your calculations of Rhat from Biogeography2.Rmd and your pencil/paper calculations.
```{r}
Rhat$Rhat[2]<-15.99        #add the value of Rhat for small, 6m replicate 1
Rhat$Rhat[3]<-19.89     #add the value of Rhat for large, 6m
Rhat$Rhat[4]<- 18.98     #add the value of Rhat for large, 2m
Rhat$Rhat[5]<-  12.36      #add the value of Rhat for small, 2m
```
Write the csv
```{r}
write.csv(Rhat,here("Data", "Outputs", "Biogeography", "rhat_complete.csv"))
```


Now summarize the data
```{r}
Rhat_summary <- Rhat %>% group_by(Size, Distance) %>% summarize(
  NRhat = length(Rhat),
  meanRhat = mean(Rhat, na.rm = T),
  sdRhat = sd(Rhat, na.rm = T), 
  semRhat = sdRhat/NRhat
)

```
###Step 3. Plot the data
Now let's look at the plot of equilibrium R as a function of island distance and size

First make some factor variables
```{r}
Rhat_summary$Distance <-as.factor(Rhat_summary$Distance)
```
Now create the plot

```{r}
ggplot(Rhat_summary, aes(Distance,meanRhat, group = Size, color = Size))+
  geom_point()+
  geom_line()+
  geom_errorbar(aes(ymax = meanRhat + semRhat, ymin = meanRhat - semRhat), width = 0.15)+
  labs(x = "Distance from mainland", y = "Mean equilibrium species richness")
```
Now save the plot

```{r}
ggsave(here("Data", "Outputs", "Biogeography","Rhat_results.png"), width = 6, height = 4)
```

