---
title: "Biogeography"
author: "Erika Barthelmess"
date: "2/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Step 1. Set up the R workspace

```{r, echo = FALSE}
rm(list = ls()) #clears the workspace for a fresh start
library(dplyr) #turns on dplyr package
library(ggplot2) #turns on ggplot2 package
library(here) #helps with file paths
```

### Step 2. Import and organize Data

#### A.
Open the master data file

```{r}
library(readr)
Data <- read.csv(here("Data", "Biogeography", "Biogeography_Data_All_Years.csv"), stringsAsFactors = T)

str(Data) #presents how the data frame is structured
```
#### B. 
Summarize the data a bit

Begin by controlling variable types a bit
```{r}
Data$Replicate<-as.factor(Data$Replicate)
```

Now summarize
```{r}
Reps<-Data %>% group_by(Size, Distance) %>% summarize(
	Replicates = length(unique(Replicate))
)
RepsByYear<-Data %>% group_by(Year, Size, Distance) %>% summarize(
   Replicates = length(unique(Replicate))
)
```

#### C. 
Now get just the data from this year
```{r}
DF <- Data %>% filter(Year == 2023)
```

Now filter the data to get each island size/dist combination

```{r}
S2M<-filter(DF, Size == "Small" & Distance == 2)
L2M<-filter(DF, Size == "Large" & Distance == 2)
S6M1<-filter(DF, Size == "Small" & Distance == 6, Replicate == "1")
S6M2<-filter(DF, Size == "Small" & Distance == 6, Replicate == "2")
L6M<- filter(DF, Size == "Large" & Distance == 6)

```
###Step 3. Perform the analysis

We run a regression of Colonization vs. Species Richness to get the regression line.  We will use data from the small, 6m island, replicate 2.  **You will need to do this again for all of the size/distance combinations, replicate 1.**

First let's look at a basic regression equation and how to structure it in R:
```{r}
IS6M2<-lm(I~R, data = S6M2) #creates the regression model for the colonization (I) vs. R for the Small (S), 6m (6M) island, replicate 2.

summary(IS6M2) #shows results
```

How would you run the regression of extinction vs. species richness?
```{r}
#what code goes here?
```


We can put both regressions together in the same block of code:
```{r}
IS6M2<-lm(I~R, data = S6M2) #creates the regression model for the colonization (I) vs. R for the Small (S), 6m (6M) island, replicate 2.

summary(IS6M2)

ES6M2 <- lm(E ~ R, data = S6M2)
summary(ES6M2)


```
Now let's plot our results using the ggplot2 package.  We want to plot I vs. R and E vs. R on the same graph.
```{r}
#Now plot together to visually find equilibrium 
ggplot(S6M2, aes(R, I))+
   geom_point(color = "blue")+
   geom_point(aes(R, E), color = "red")+
   geom_abline(data=ES6M2, intercept = ES6M2$coefficients[1], 
               slope = ES6M2$coefficients[2], color = "red")+
   geom_abline(data=IS6M2, intercept = IS6M2$coefficients[1], 
               slope = IS6M2$coefficients[2], color = "blue")+
	geom_vline(xintercept = 2.96, linetype = "dashed")+
  labs(x="R (species richness)", y = "I or E", title = "Small Island, 6 m, replicate 2")
  
```

And now we can save our plot for later (like to insert in your homework!)
```{r, echo = False}
#now save your result
ggsave(here("Data", "Outputs","Biogeography","Small 6M replicate 2equilibrium.png"), width = 6, height = 4,device = "png")

```

  





