---
title: "Biogeography"
author: "Erika Barthelmess"
date: "2/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Step 1. Set up the R workspace

```{r, echo = FALSE}
rm(list = ls()) #clears the workspace for a fresh start
library(dplyr) #turns on dplyr package
library(ggplot2) #turns on ggplot2 package
library(here) #helps with file paths
```

### Step 2. Import and organize Data

#### A.
Open the master data file

```{r}
library(readr)
Data <- read.csv(here("Data", "Biogeography", "Biogeography_Data_All_Years.csv"), stringsAsFactors = T)

str(Data) #presents how the data frame is structured
```
#### B. 
Summarize the data a bit

Begin by controlling variable types a bit
```{r}
Data$Replicate<-as.factor(Data$Replicate)
```

#### C. 
Now get just the data from this year
```{r}
DF <- Data %>% filter(Year == 2023)
```

Now filter the data to get each island size/dist combination

```{r}
S2M<-filter(DF, Size == "Small" & Distance == 2)
L2M<-filter(DF, Size == "Large" & Distance == 2)
S6M1<-filter(DF, Size == "Small" & Distance == 6, Replicate == "1")
S6M2<-filter(DF, Size == "Small" & Distance == 6, Replicate == "2")
L6M<- filter(DF, Size == "Large" & Distance == 6)

```
###Step 3. Perform the analysis
We did the work for the Small, 6m island, replicate 2. Now it is your turn to set up the code for the remaining island size/distance combinations.
**You will need to do this again for all of the size/distance combinations, replicate 1.**

**Use the code displayed in lines 61 - 80 as a model to fill in the code blocks that appear below.**

```{r}
IS6M2<-lm(I~R, data = S6M2) #creates the regression model for the colonization (I) vs. R for the Small (S), 6m (6M) island, replicate 2.

summary(IS6M2)

ES6M2 <- lm(E ~ R, data = S6M2)
summary(ES6M2)

#Now plot together to visually find equilibrium 
ggplot(S6M2, aes(R, I))+
   geom_point(color = "blue")+
   geom_point(aes(R, E), color = "red")+
   geom_abline(data=ES6M2, intercept = ES6M2$coefficients[1], 
               slope = ES6M2$coefficients[2], color = "red")+
   geom_abline(data=IS6M2, intercept = IS6M2$coefficients[1], 
               slope = IS6M2$coefficients[2], color = "blue")+
  labs(x="R (species richness)", y = "I or E", title = "Small Island, 6 m, replicate 2")
  
```

### Small, 2m
Fill in the following code chunk to get the I and E equations for the small, 2m island treatment. Modify the plot so that the title reads "Small, 2m island"

```{r}

```

Now fill in the following code chunk to save your plot to the Biogeography folder in the Outputs folder. Name the plot "small_2m.png"
```{r}

```

### Small, 6m replicate 1
Fill in the following code chunk to get the I and E equations for the small, 6m island replicate 1 treatement. Modify the plot so that the title reads "Small, 6m island replicate 1"

```{r}

```

Now fill in the following code chunk to save your plot to the Biogeography folder in the Outputs folder. Name the plot "small_6m_rep1.png"
```{r}

```

### Large, 2m replicate
Fill in the following code chunk to get the I and E equations for the large, 2m island. Modify the plot so that the title reads "Large 2m island"

```{r}

```

Now fill in the following code chunk to save your plot to the Biogeography folder in the Outputs folder. Name the plot "large_2m.png"
```{r}

```

### Large, 6m replicate
Fill in the following code chunk to get the I and E equations for the large, 6m island. Modify the plot so that the title reads "Large 6m island"

```{r}

```

Now fill in the following code chunk to save your plot to the Biogeography folder in the Outputs folder. Name the plot "large_6m.png"
```{r}

```


Great! Next you need to solve for equilibrium R for each of these experiments.